---
title: "Education and Political Approval"
author: "Alec Rogers"
format: html
execute: 
  echo: false
---


```{r setup}
#| message: false
library(MASS)
library(tidyverse)
library(primer.data)
library(broom)
library(easystats)

```

Political ideology relates to presidential approval. We use 2020 data from the CES survey to explore this relationship among American voters.Since our data come from only 2020 and from a survey, applicability of our findings is limited to other time periods and populations that may not respond to political surveys.


```{r}
# create the data set
x <- ces |>
filter(year == 2020) |>
dplyr::select(approval, education, ideology) |>
drop_na() |>
slice_sample(n = 2000)

```


```{r}
x <- ces |>
  filter(year == 2020) |>
  select(approval, ideology, education) |>
  drop_na() |> 
  filter(! ideology %in% "Not Sure") |> 
  mutate(ideology = fct_drop(ideology))

```

$$P(Y \leq k) = \frac{1}{1 + e^{-(\alpha_k - \beta_1 \cdot \text{ideology} - \beta_2 \cdot \text{education})}}$$
x
with $Y \sim \text{Ordinal}(\boldsymbol{\rho})$ where $\boldsymbol{\rho} = (\rho_1, \rho_2, \ldots, \rho_K)$ are derived from the cumulative probabilities above.


```{r}
fit_approval = polr(formula = approval ~ ideology + education, data = x, Hess = TRUE)
```


```{r}
pp <- check_predictions(fit_approval)

pp |>
 as.data.frame() |>
 str(max.level = 1)
```

```{r}
#| cache: true
approval_table <- fit_approval |>
tidy(conf.int = TRUE)

```

```{r}
approval_table |> 
  select(term, estimate, conf.low, conf.high) |> 
  gt() |> 
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "Lower CI",
    conf.high = "Upper CI"
  )
```
